---
title: "Graph clustering with the Stochastic block model (using Daudin's ICL)"
author: "Martin Metodiev"
output:
  pdf_document:
    toc: no
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

```{css, echo=FALSE}
p.caption {
  font-style: italic;
}
```

```{r setup-packages,message=FALSE, include=FALSE, echo=FALSE}
library(Matrix) # sparse matrix
library(FixedPoint) # Fixed point solution for the variational M-step
library(cluster) # Clustering to find the initial value
source("algorithms.R")
source("classes.R")
set.seed(2134)

```


# Daudin's Paper

The following is a brief project with the goal to implement Daudin's ICL:

```{=tex}
\begin{equation}
\label{eq:sbm}
\begin{aligned}
ICL(m_Q)&=\max_\theta\log\mathcal{L}(\mathcal{X},\tilde{\mathcal{Z}}|\theta,m_Q)\\&-\frac{1}{2}\times\frac{Q(Q+1)}{2}\log\frac{n(n-1)}{2}-\frac{Q-1}{2}\log(n)
\end{aligned}
\end{equation}
```

## Simulation scenario

The following simulation scenario was inspired from a vignette from the greed package. However, none of this package's code is used. All new functions used were written by myself.

We begin by simulating from a hierarchically structured SBM model, with 2 large clusters, each composed of 3 smaller clusters with higher connection probabilities, making a total of 6 clusters.

```{r, fig.show='hold'}
N  <- 400           # Number of node
K  <- 6             # Number of cluster
pi <- rep(1/K,K)    # Clusters proportions 
lambda   <- 0.1     # Building the connectivity matrix template
lambda_o <- 0.01
Ks <- 3
mu <- bdiag(lapply(1:(K/Ks), function(k){
  matrix(lambda_o,Ks,Ks)+diag(rep(lambda,Ks))}))+0.001
sbm <- sim_sbm(N,pi,mu) # Simulation
```

The method finds a very good solution. Model selection is done via Daudin's ICL as described above. Note that convergence thresholds are set to be quite small (1e-3), so large errors can (and do) occur, especially when the number of classes Q is large. In fact, setting Q<=6 is recommended.

```{r, fig.show='hold'}
sol = var_bayes_model_selection(sbm$x, Q=3)
table(sol@cl, sbm$cl)
```
