---
title: "Graph clustering with the Stochastic block model (using Daudin's ICL)"
author: "Martin Metodiev"
output:
  pdf_document:
    toc: yes
    toc_depth: '1'
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

```{css, echo=FALSE}
p.caption {
  font-style: italic;
}
```


Loads packages and set a future plan for parallel processing.
```{r setup-packages,message=FALSE, include=TRUE}
library(Matrix) # sparse matrix
library(FixedPoint) # Fixed point solution for the variational M-step
library(cluster) # Clustering to find the initial value
source("sbm_prototype.R")
source("classes.R")
set.seed(2134)

```


# Daudin's Paper

The following is a brief project with the goal to implement Daudin's ICL:

```{=tex}
\begin{equation}
\label{eq:sbm}
\begin{aligned}
\pi&\sim \textrm{Dirichlet}_K(\alpha),\\
\theta_{k,l} & \sim \textrm{Beta}(a_0, b_0), \\
Z_i&\sim \mathcal{M}(1,\pi),\\
\forall (i,j), \quad x_{ij} \mid Z_{ik}Z_{jl}=1& \sim \mathcal{B}(\theta_{k,l}).
\end{aligned}
\end{equation}
```

## Simulation scenario

The following simulation is copied from a vignette from the greed package. However, none of its code is used.

We begin by simulating from a hierarchically structured SBM model, with 2 large clusters, each composed of 3 smaller clusters with higher connection probabilities, making a total of 6 clusters. Greed comes shipped with simulation function for the different generative models it handle and we will take advantage of the ``rsbm()` function to simulate an SBM graph with 6 clusters and 400 nodes:
```{r, fig.show='hold'}
N  <- 400           # Number of node
K  <- 6             # Number of cluster
pi <- rep(1/K,K)    # Clusters proportions 
lambda   <- 0.1     # Building the connectivity matrix template
lambda_o <- 0.01
Ks <- 3
mu <- bdiag(lapply(1:(K/Ks), function(k){
  matrix(lambda_o,Ks,Ks)+diag(rep(lambda,Ks))}))+0.001
sbm <- sim_sbm(N,pi,mu) # Simulation
```

```{r, fig.show='hold'}
d_test = var_bayes_model_selection(sbm$x, Q=3)
table(d_test@cl, sbm$cl)
